Planning to Save Energy: How Information Format Affects Accuracy

### 2024-12-07 manuscript/manuscript.qmd
update results section. And some formating improvements.

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
author:
  - name: Thomas E. Gorman
    url: https://tegorman13.github.io/
    orcid: "0000-0001-5366-5442"
    affiliations: 
      - ref: cclab
      - ref: clara
    attributes: 
      corresponding: true
  - name: Torsten Reimer
    url: https://web.ics.purdue.edu/~treimer/
    orcid: 0000-0002-7419-0076
    affiliations:
      - ref: cclab
      - ref: clara
  - name: Juan Pablo Loaiza Ramirez
    url: https://www.cla.purdue.edu/directory/profiles/loaiza-ramirez,juan-pablo.html
    orcid: 0000-0001-9663-0522
    affiliations:
      - ref: cclab
      - ref: clara
  - name: Hayden Barber
    url: https://www.sdstate.edu/directory/hayden-barber
    orcid: 0000-0002-3465-8615
    affiliations:
      - ref: sdsu
affiliations:
  - id: cclab
    name: Communication and Cognition Lab, Purdue University
    city: West Lafayette
    region: IN
    country: USA
    postal-code: 47907
    url: https://web.ics.purdue.edu/~treimer/
  - id: clara
    name: College of Liberal Arts Research Academy
    department: Purdue University
    city: West Lafayette
    region: IN
    country: USA
    postal-code: 47907
    url: https://cla.purdue.edu/about/college-initiatives/research-academy/
  - id: sdsu
    name: School of Communication & Journalism, South Dakota State University
    city: Brookings
    region: SD
    country: USA
    url: https://www.sdstate.edu/directory/hayden-barber
format:
    # html: 
    #     grid:
    #         sidebar-width: 220px
    #         body-width: 1200px
    #         margin-width: 170px
    #         gutter-width: 1.0rem
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        code-fold: true
        code-summary: "Show the code"
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    hugo-md:
        include: true
        echo: false
        html-math-method: mathjax
        output-file: manuscript_hugo.md
    gfm:
        echo: true
        output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
# format-links:   
#     - html
#     - gfm
#     - format: hikmah-manuscript-pdf
#       text: PDF version
#       icon: file-pdf
date: 2024-12-03
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption (Fischer, 2008)

@@

- Canfield et al. (2017) found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading. However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
- Canfield et al.'s (2017) findings on preferences for historical use information and the impact of neighbor comparisons
- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements (Vessey, 1991)
- alignment can reduce cognitive load and enhance accuracy in planning (Shah & Freedman, 2009)
- Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.

Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021 . This challenge partly arises from difficulties in converting information across numerical formats, impeding the development of precise energy reduction plans @reimerNumericCommunicationRisk2015. Prior research by @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that presenting energy information in tabular formats enhances comprehension relative to graphs.

The way numerical information is presented can significantly affect how individuals process and use that information (Reimer et al., 2015). The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to (Gigerenzer & Edwards, 2003; Reimer et al., 2015). Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.

Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities (Gigerenzer & Hoffrage, 1995; Hoffrage et al., 2000). In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.

@@

```{r}

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

@@

```{r}
#| label: tbl-s1-agg
#| caption: "Study 1: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

@@

For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error.

```{r}
#| label: tbl-s1-prop
#| caption: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

@@

```{r}
#| label: fig-s1-plot
#| caption: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

@@

```{r}
#| label: tbl-s1-ord
#| caption: "Study 1: Odds ratios of for group comparisons. "

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 2), class = "Intercept"), 
                prior(normal(0, 1.5), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)

#summary(ordinal_model_s1)

# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
#     filter(stringr::str_detect(Parameter, "b_")) |> 
#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
#     mutate(across(where(is.numeric), round, 3)) |> 
#     kable(booktabs = TRUE)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, .category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

# Convert log-odds to odds ratios
posterior_samples <- as.data.frame(ordinal_model_s1)
odds_ratios <- data.frame(
  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
)

# Calculate summary statistics
odds_ratio_summary <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh"),
    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
                    mean(odds_ratios$USD_vs_kWh)),
    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
                quantile(odds_ratios$USD_vs_kWh, 0.025)),
    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
                quantile(odds_ratios$USD_vs_kWh, 0.975))
)

odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
# |comparison        | odds_ratio| ci_lower| ci_upper|
# |:-----------------|----------:|--------:|--------:|
# |Percentage vs kWh |        2.3|     0.62|      6.2|
# |USD vs kWh        |       14.1|     4.08|     38.0|

#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")

```{r}
#| fig.width: 11
#| fig.height: 7

bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])

# Plot predicted probabilities
ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                    position = position_dodge(width = 0.9), width = 0.2) +
    labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
    ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
    theme_minimal()

@@

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
    # scale_x_continuous("y", breaks = 1:7) +
    # scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
    ggtitle("Data with posterior predictions",
            subtitle = "S1. red bars are empirical data, black points are model predictions and credible intervals") +
    labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
    theme_minimal() +
    theme(legend.background = element_blank() #,legend.position = c(.9, .75)
    )

@@

```{r}
#| label: tbl-s2-agg
#| caption: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# overall pct of subjects who matched their goal
s2_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Deviation From Goal' = median(mean_pct_change),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

@@

```{r}
#| label: tbl-s2-prop
#| caption: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

@@

```{r}
#| label: fig-s2-plot
#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

@@

```{r}
#| label: fig-s2-plot2
#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
#| fig.width: 11
#| fig.height: 9

@@

@@

```{r}
#| label: fig-s2-indv2
#| caption: "Study 2: Respones patterns for a subset of individiual participants. Black points are participant responses, red points are the state average, and blue points are the family average. The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
#| fig.width: 11
#| fig.height: 10

>>>>

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
format:
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    html: 
      echo: true
      code-fold: show
      output-file: manuscript-code.html
    hugo-md:
        include: false
        echo: false
        html-math-method: mathjax
        output-file: manuscript_hugo.md
    gfm:
        echo: true
        output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
format-links:   
    - text: HTML
      href: manuscript.html
      icon: file-minus
    - text: HTML (with code)
      href: manuscript-code.html
      icon: file-plus
    - format: hikmah-manuscript-pdf
      text: PDF version
      icon: file-pdf
    - hikmah-manuscript-docx
    - gfm
date: 2024-12-07
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption [@fischerFeedbackHouseholdElectricity2008]

@@

- @canfieldPerceptionsElectricityuseCommunications2017 found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading. However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
- @canfieldPerceptionsElectricityuseCommunications2017's findings on preferences for historical use information and the impact of neighbor comparisons
- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991]
- alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011]
- Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.

Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021. This challenge partly arises from difficulties in converting information across numerical formats, impeding the development of precise energy reduction plans @reimerNumericCommunicationRisk2015. Prior research by @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that presenting energy information in tabular formats enhances comprehension relative to graphs.

The way numerical information is presented can significantly affect how individuals process and use that information [@reimerNumericCommunicationRisk2015] . The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015]. Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.

Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities [@gigerenzerHowImproveBayesian1995 @hoffrageCommunicatingStatisticalInformation2000]. In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.

@@

```{r}
pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

@@

```{r}
#| label: tbl-s1-agg
#| tbl-cap: "Study 1: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

@@

For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error. The current analysis employs a cumulative ordinal regression model, implemented via a Bayesian hierarchical framework [@burknerBrmsPackageBayesian2017] . This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category. All analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].

```{r}
#| label: tbl-s1-prop
#| tbl-cap: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

@@

see @fig-s1-plot

```{r}
#| label: fig-s1-plot
#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

@@

```{r}
#| label: tbl-s1-ord
##| tbl-cap: "Study 1: Ordinal Regression Model Results."

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 5000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)

#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

# as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, Category=.category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

#plot(conditional_effects(ordinal_model_s1),points=TRUE)

# or_table |>
#       rownames_to_column(var = "Term") |>
#       filter(!stringr::str_detect(Term, "Intercept")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))

# or_table <- data.frame(exp(fixef(ordinal_model_s1)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")
```

::: {#tbl-s1-or}
|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-1.94    |-3.19    |-0.74    |1.00 |
|Intercept[2]       |0.20     |-1.02    |1.42     |0.63 |
|Intercept[3]       |4.36     |3.11     |5.65     |1.00 |
|refClassPercentage |0.85     |-0.54    |2.23     |0.88 |
|refClassUSD        |2.72     |1.41     |4.06     |1.00 |

**Experiment 1**: Ordinal Regression results.

:::

::: {#tbl-s1-epred}
|Term               |Estimate |Est.Error |Q2.5 |Q97.5 |
|:------------------|:--------|:---------|:----|:-----|
|refClassPercentage |2.3      |2         |0.58 |9.3   |
|refClassUSD        |15.2     |2         |4.11 |58.1  |

**Experiment 1**: Odds ratios for group comparisons.

:::

The ordinal model is parameterized with thresholds (intercepts), and positive coefficients can indicate that it is more difficult to achieve higher accuracy categories in the USD condition. 
The model output suggests that, compared to the kWh condition, the USD condition shows a positive coefficient (Estimate = 2.72, 95% CI: 1.41 to 4.06) for the ordinal outcome. At least to me, this positive coefficient appears to indicate that, relative to the kWh reference class, participants in the USD condition are more likely to fall into higher numerical categories of the dependent variable coding. However, because the dependent variable is ordered from best (Exact match) to worst (Over 15% error), care is needed in interpretation. The Percentage condition coefficient (Estimate = 0.85, 95% CI: -0.54 to 2.23) is more uncertain, with its credible interval overlapping zero.  Posterior predictive checks (Figure 3) showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-plot2).

```{r}
#| label: fig-s1-plot2
#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
#| fig.width: 11
#| fig.height: 7

#bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])

# Plot predicted probabilities
# ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = Category)) +
#     geom_bar(stat = "identity", position = position_dodge()) +
#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
#                     position = position_dodge(width = 0.9), width = 0.2) +
#     labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
#     ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
#     theme_minimal()

@@

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

@@

```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.001)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# overall pct of subjects who matched their goal
s2_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Abs. Deviation From Goal' = median(mean_pct_change),
    'Log Deviation' = (median(mean_log_abs_error)),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

# s2_agg1 |> group_by(id,refClass,calc) |> 
#     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
#     summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
#     mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
#     n_accuracy=first(n_accuracy))

# s2_agg4 %>% ggplot(aes(x=refClass,y=mean_log_abs_error,fill=refClass)) + 
#   stat_summary(fun=mean, geom="bar") +
#   stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
#   geom_jitter(alpha = 0.2, width = 0.2, height = 0)

# s2_agg4 |> 
#   ggplot(aes(x = mean_log_abs_error,fill = refClass)) +
#   geom_density(alpha = 0.7)

# library(ggridges)
# s2_agg1 |>  ggplot(aes(y = refClass, x=log_abs_error,fill=refClass,col=refClass)) +
#     geom_density_ridges(alpha = 0.5) +
#     geom_vline(aes(xintercept=(log_abs_error)))

@@

```{r}
#| label: tbl-s2-prop
#| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages). The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."

@@

```{r}
#| label: fig-s2-plot
#| tbl-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."

@@

```{r}
#| label: fig-s2-plot2
#| fig-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.  A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
#| fig.width: 11
#| fig.height: 9

@@

```{r}
#| label: tbl-s2-ord
##| tbl-cap: "Study 2: Ordinal Regression Model Results."

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
#summary(ordinal_model_s2_logit)

# as.data.frame(describe_posterior(ordinal_model_s2_logit, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
#   kable(escape = FALSE, booktabs = TRUE, align = c("l"), row.names = FALSE)

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
 # group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
  group_by("Reference Class"=refClass, Category=.category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

#pred_summary_s2 |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)

# Create a new data frame with the predictor values
# new_data <- expand.grid(
#   refClass = c("kWh", "Percentage", "USD"),
#   id = unique(s2_agg1$id)[1],      # Take first id
#   state = unique(s2_agg1$state)[1], # Take first state,
#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
# )

# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
# or_table |>
#   rownames_to_column(var = "Term") |>
#   filter(!stringr::str_detect(Term, "Intercept")) |>
#   as_flextable() |>
#   set_header_labels(values = list(
#     OR = "Odds Ratio",
#     `2.5 %` = "Lower 95% CI",
#     `97.5 %` = "Upper 95% CI"
#   )) |>
#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
#   align(align = 'center', part = 'all') |>
#   fontsize(size = 10, part = "all")

::: {#tbl-s2-or}
|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
|:------------------|:--------|:--------|:--------|:----|
|Intercept[1]       |-2.13    |-3.39    |-0.86    |1.00 |
|Intercept[2]       |-0.62    |-1.89    |0.63     |0.84 |
|Intercept[3]       |3.15     |1.88     |4.42     |1.00 |
|refClassPercentage |0.83     |-0.64    |2.33     |0.87 |
|refClassUSD        |1.87     |0.31     |3.35     |0.99 |
|roundedRounded     |-0.66    |-1.01    |-0.31    |1.00 |
|pct_goal15%        |-0.44    |-0.79    |-0.10    |0.99 |

**Experiment 2.** Ordinal Regression Model Results.

:::

::: {#tbl-s2-epred}
|comparison           | odds_ratio| ci_lower| ci_upper|
|:--------------------|----------:|--------:|--------:|
|Percentage vs kWh    |       3.02|     0.53|    10.31|
|USD vs kWh           |       8.80|     1.37|    28.42|
|Rounded vs Not       |       0.53|     0.36|     0.73|
|15% Goal vs 10% Goal |       0.66|     0.45|     0.91|

**Experiment 2.** Odds ratios for group comparisons.

:::

```{r}
#| label: fig-s2-plot3

# ggplot(pred_summary_s2, 
#         aes(x = .category, y = mean_prob, fill = refClass )) +
#     geom_col(position = position_dodge(width = 0.8)) +
#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
#                     position = position_dodge(width = 0.8)) +
#     labs(x = "Reference Class", 
#         y = "Predicted Probability",
#         fill = "Accuracy Level",
#         title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
#     scale_fill_brewer(palette = "RdYlBu") +
#     theme_minimal()

pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
#| label: fig-s2-indv2
#| fig-cap: "Study 2: Respones patterns for a subset of individiual participants. Black points are participant responses, red points are the state average, and blue points are the family average. The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
#| fig.width: 11
#| fig.height: 10
